Formal Verification Framework Based on Symbolic Execution for Smart Contract

### 2023-12-29 papers/static-checker/ecoop2024/ecoop2024.tex
kolbri

@@

\begin{lstlisting}[float,captionpos=b,caption={Kolibri oracle contract},label={lst:kolibri-oracle-contract},numbers=left]
parameter
  ( or
    ( or
      ( unit %default )
      ( contract %getXtzUsdRate nat ) )
    ( or
      ( pair %getXtzUsdRate_callback string 
      ( pair timestamp nat ) )
      ( or
        ( address %setGovernorContract )
        ( nat %setMaxDataDelaySec ) ) ) ) ;
storage
  ( pair
    ( pair ( option %clientCallback address ) 
          ( address %governorContractAddress ) )
    ( pair ( address %harbingerContractAddress ) 
          ( pair ( nat %maxDataDelaySec ) ( int %state ) ) ) ) ;
\end{lstlisting}

\section {Conclusion}
\section {Appendix}
\end{document}

>>>>

The ability to generate fail conditions is a potent feature of our static checker, particularly in verifying access control properties across various smart contracts. Specifically, in numerous smart contract configurations, certain entrypoints are designated for invocation only by authorized entities, often exemplified by the contract owner. Consequently, these entrypoints must result in failure if the callers lack the necessary authorization. This crucial access control property is verified by examining the fail conditions of a smart contract, ensuring that unauthorized callers trigger the intended failures in the system.

@@

\begin{lstlisting}[float,captionpos=b,caption={Kolibri oracle contract},label={lst:kolibri-oracle-contract},numbers=left]
parameter
 (or
   (or
     (unit %default )
     (contract nat %getXtzUsdRate ))
   (or
     (pair string (pair timestamp nat) %getXtzUsdRate_callback)
     (or
       (address %setGovernorContract)
       (nat %setMaxDataDelaySec)))) ;
storage
 (pair
   (pair (option %clientCallback address) 
         (address %governorContractAddress))
   (pair (address %harbingerContractAddress) 
         (pair (nat %maxDataDelaySec) (int %state)))) ;
\end{lstlisting}

Another critical access control property pertains to the resettable nature of \lstinline/governorContractAddress/ and \lstinline/maxDataDelaySec/. While these parameters are designed to be reset, such actions should exclusively be carried out by the governor. Maintaining this restriction is paramount to prevent loss of control over the contract in the event of unauthorized resets. In practical terms, the entrypoints \lstinline/setGovernorContract/ and \lstinline/setMaxDataDelaySec/ must only be invocable by the governor. Any attempts by unauthorized callers to make these calls should result in failure to ensure the integrity of access control and prevent potential misuse of these critical parameters.

The Kolibri oracle contract maintains a straightforward state machine comprising two states: (1) \lstinline/IDLE/ and (2) \lstinline/WAITING_FOR_HARBINGER/. From the \lstinline/IDLE/ state, the oracle can exclusively transition to the \lstinline/WAITING_FOR_HARBINGER/ state. The \lstinline/getXtzUsdPrice/ entrypoint is permissible only when the oracle is in the \lstinline/IDLE/ state, and invoking this entrypoint transitions the oracle to the \lstinline/WAITING_FOR_HARBINGER/ state. Notably, in the \lstinline/IDLE/ state, the value of \lstinline/clientCallback/ is always set to none. Conversely, from the \lstinline/WAITING_FOR_HARBINGER/ state, the oracle can only transition back to the \lstinline/IDLE/ state. The \lstinline/getXtzUsdPrice_callback/ entrypoint is allowable solely when the oracle is in the \lstinline/WAITING_FOR_HARBINGER/ state, and invoking this entrypoint results in the oracle transitioning to the \lstinline/IDLE/ state. In the \lstinline/WAITING_FOR_HARBINGER/ state, the value of \lstinline/clientCallback/ is always set to some.

Several properties must be verified to ensure the correct functioning of the state machine. The initial property states that \lstinline/getXtzUsdPrice/ can only be invoked when in the \lstinline/IDLE/ state, while conversely, \lstinline/getXtzUsdPrice_callback/ can only be invoked when in the \lstinline/WAITING_FOR_HARBINGER/ state. The validation of this property involves examining the fail conditions of these two entrypoints.

The subsequent property dictates that if \lstinline/getXtzUsdPrice/ is successfully executed, the state must transition to \lstinline/WAITING_FOR_HARBINGER/, and the parameters passed along with the call need to be stored in the \lstinline/clientCallback/ attribute of the storage. This property can be detailed as shown in Listing \ref{lst:kolibri-contract-specification}.

The following property asserts that if \lstinline/getXtzUsdPrice_callback/ is successful, the state should revert to \lstinline/IDLE/, and there must be a transaction transferring the data to the client at the \lstinline/clientCallback/ attribute. This property is specified in Listing \ref{lst:kolibri-contract-specification}.

The final property concerns the entrypoint relation between \lstinline/getXtzUsdPrice/ and \lstinline/getXtzUsdPrice_callback/ as depicted in Figure \ref{fig:kolibri-oracle-emtrypoint-relations}. From the \lstinline/getXtzUsdPrice/ entrypoint, the contract can only be called with \lstinline/getXtzUsdPrice_callback/ and then reverted. Invoking the \lstinline/getXtzUsdPrice/ entrypoint after another call or invoking the \lstinline/getXtzUsdPrice_callback/ entrypoint again is not allowed. This entrypoint relation property is specified as an entrypoint relation in Listing \ref{}.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{kolibri}
    \caption{Kolibri oracle emtrypoint relations}
    \label{fig:kolibri-oracle-emtrypoint-relations}
\end{figure}

\begin{lstlisting}[float,captionpos=b,caption={Kolibri oracle contract specification},label={lst:kolibri-contract-specification},numbers=left]
mcontract Auction = spec 
  storage := (Pair (Pair (client: option address) 
     (governor: address))
     (Pair (harbinger: address) 
           (Pair (maxDelay: nat) (state: int))))
  entrypoint %getXtzUsdRate
    code := {...}
    parameter := (c : contract nat) 
    pre-condition := (Amount = 0) | (state = 0)
    post-condition := (state = 1) | 
       (client = Some Sender) ;
  entrypoint %getXtzUsdRate_callback
    code := {...}
    parameter := Pair (s: string) 
       (Pair (t: timestamp) (n: nat))
    pre-condition :=  state = 1 |
       Sender = harbinger  |
       s = 'XTZ-USD' |
       Now - t <= 30 |
       client = Some a |
       get_contract (nat, a) = Some c    
    post-condition := state = 0 |
       post_client = None |
       transfer(n, 0, c) ;
  entrypoint %setGovernorContract
    code := {...}
    parameter := (gover : address) 
    pre-condition :=  (Sender = governor)                  
    post-condition := (governor = gover) |
       (post_maxDelay = maxDelay) |
       (post_harbinger = harbinger) ;
  entrypoint %setMaxDataDelaySec
    code := {...}
    parameter := (delay: nat)
    pre-condition := (Sender = governor)             
    post-condition := (maxDelay = delay) |
       (post_client = client) |
       (post_harbinger = harbinger) ;
  (%getXtzUsdRate -> %getXtzUsdRate_callback) with
     (Sender = harbinger)
     && (s = 'XTZ-USD') 
     && (Now - t <= 1800)  | 
  (%getXtzUsdRate_callback  -> %getXtzUsdRate) with 
    (Amount = 0) 
  not (%getXtzUsdRate -> %getXtzUsdRate) |
  not (%getXtzUsdRate_callback -> %getXtzUsdRate_callback)
\end{lstlisting}
\section {Conclusion}
\section {Appendix}
\end{document}
